---
title: "Lab 8 - Linear Regression in R"
author: "Rachel Behm"
date: "November 19, 2018"
output: html_document
---
```{r, include=FALSE}
#load in packages and csv file
library(tidyverse)

salmon <- read_csv("salmon_imports_simplified.csv")
```

##Exploratory scatterplot

If we’re going to use linear regression, we should have reason to think that a linear relationship exists between
variables. 

So go ahead and look at the data (year on x-axis, imports on y-axis).

__a.__ Make an exploratory scatterplot in ggplot (recall: geom_point). Do not finalize.

__b.__ As a comment in the code chunk, answer: Overall, does it look like describing the relationship between
year and fresh salmon imports using linear regression makes sense? Do you have any concerns about
using linear regression to describe the relationship?

```{r}
salmon_scatter<- ggplot(salmon, aes(x=year, y=salmon_fresh)) +
  geom_point()
salmon_scatter
```

##Linear regression in R

**The general syntax for linear regression in R is:**

model_name <- lm(y_variable ~ x_variable, data = df_name)

So, if I have a data frame called ‘salmon’ containing variables ‘year’ and ‘salmon_fresh,’ __I would use__:

salmon_model <- lm(salmon_fresh ~ year, data = salmon)

_________________________________________________________________________________________________________________
__a.__ Perform linear regression to describe the trend in total volume of fresh salmon imported to the U.S.
from 1990 - 2017.

__b.__ Call the model name to see the intercept and slope for the model. Recall:
y = 0 + 1x + 

__c.__ Write out the model equation, including the actual variables and coefficients instead of  and x/y. For
example (these are not the actual model coefficients you’ll find):
Imports(tons) = −400 + 25(Y ear)

__d.__ Think about the model equation that you found to describe trends in salmon imports. In your markdown
document, answer the following in 1-sentence each:
• What does the slope mean in the context of this model?
• What does the y-intercept mean in the context of this model? Why is that concerning? What does this
mean about extrapolating this model for past values?

```{r}
#syntax for making a linear regression with this data
salmon_model <- lm(salmon_fresh ~ year, data = salmon)
salmon_model
```

##Model diagnostics
Use plot(model_name) to view model diagnostics in the ‘Plots’ tab 

(press Enter in the Console to contine to subsequent diagnostic plots). 

__Explore the diagnostic plots to answer the following:__
• Do residuals appear normally distributed?
• Any concerns about heteroscedasticity or outliers?

To view all four diagnostic plots at once (and have them appear in your knitted document), you can use the
following:

par(mfrow = c(2,2))
plot(model_name)

```{r}
#view model diagnostics in plots tab
#plot(salmon_model)

#to make all appear together and save space
par(mfrow= c(2,2))
plot(salmon_model)
```

